\include{preamble}

\setlength{\textheight}{23cm} %{23cm}
\setlength{\topmargin}{-2cm}
\setlength{\textwidth}{17.5cm} \setlength{\oddsidemargin}{-0.5cm}
\setlength{\evensidemargin}{-0.5cm}

\setlength{\parindent}{0pt}

\newcommand{\gap}{\vspace{5pt}}
\newcommand{\epc}{\hspace{1pc}}

\newcommand{\onebld}{{\bf 1}}
\newcommand{\wt}{\widetilde}
\newcommand{\wh}{\widehat}

\newcommand{\E}{{\rm I\!E}}
\newcommand{\IP}{{\rm I\!P}}
\newcommand{\D}{{\rm I\!D}}
\newcommand{\pmat}[1]{\begin{pmatrix} #1 \end{pmatrix}}
\newcommand{\us}[1]{{\color{black}#1}}
\newcommand{\ssbs}[1]{{\color{blue}#1}}
\bibliographystyle{plain}

%% Semih's additions %%
\usepackage{mathtools}
%% end of semih's additions %%


\title{\bf Taming the Duck: Can Stochastic Programming Help?}
\date{}
%\author{Harsha Gangammanavar\\Engineering Management, Information, and Systems\\Southern Methodist University, \\ Dallas, TX 75275}

\begin{document}
\pagenumbering{gobble}
\maketitle

\vspace*{-1.3cm}

\section{Formulations}

\subsection{Decision Variables and Parameters}

Let $\mathcal{G}$ represent the set of generators, and $\mathcal{T}$ denote the set of discretized time periods. The set of buses is denoted with $\mathcal{B}$ and the set of lines is denoted with $\mathcal{L}$. The generators that are located at a particular bus $j$ are represented with the set $\mathcal{G}_j$. The following variables are defined for each generator $g \in \mathcal{G}$, and period $t \in \mathcal{T}$:

\begin{description} \setlength{\itemsep}{-1pt}
\item{$x_{gt}$:} 1 if generator $g$ is operational at $t$, 0 otherwise,
\item{$s_{gt}$:} 1 if generator $g$ becomes online at $t$, 0 otherwise,
\item{$z_{gt}$:} 1 if generator $g$ becomes offline at $t$, 0 otherwise,
\item{$G_{gt}$:} Production level of generator $g$ at $t$, 
\end{description}

For each bus $j\in \mathcal{B}$ and period $t \in \mathcal{T}$, we define the following variables: 
\begin{description}
\item{$\theta_{jt}$:} Voltage angle at bus $j$ at $t$,
\item{$L_{jt}$:} Amount of load shed at bus $j $ at $t$.
\item{$O_{jt}$:} Amount of over-generation at bus $j$ at $t$.
\end{description}
The latter two variables are penalized in the objective with the coefficients $L^{penalty}$ and $O^{penalty}$, respectively. 

Each generator $g \in \mathcal{G}$ is characterized by the following parameters:
\begin{description} \setlength{\itemsep}{-1pt}
\item{$G_g^{\max}$:} Maximum generation capacity, 
\item{$G_g^{\min}$: } Minimum generation requirement when the generator is online, 
\item{$\Delta G_g^{\max}$: } Ramp up limit, 
\item{$\Delta G_g^{\min}$:} Ramp down limit,
\item{$UT_g$:} Minimum required uptime before the generator can become offline, 
\item{$DT_g$:} Minimum required downtime before the generator can become online,
\item{$c_g^{gen}$:} Generation cost,
\item{$c_g^{start}$:} Start up cost,
\item{$c_g^{noload}$:} No load cost.
\end{description}

The problem is defined over a topology that is defined by the following characteristics:
\begin{description} \setlength{\itemsep}{-1pt}
\item{$F^{\min}_{ij,t}$ / $F^{\max}_{ij,t}$:} Flow lower / upper limits over line $(i,j) \in \mathcal{L}$ at $t$,
\item{$B_{ij}$:} Susceptance of line $(i,j) \in \mathcal{L}$.
\end{description}

Finally, each bus $j\in \mathcal{B}$ faces a demand $D_{jt}$ at period $t \in \mathcal{T}$. 


\subsection{Unit Commitment}

\begin{subequations}
\begin{align}
\begin{split}
\min \quad \sum_{g \in \mathcal{G}} \sum_{ t \in \mathcal{T} } \left(c^{gen}_{g} G_{gt} + c^{start}_g s_{gt} + c^{noload}_g x_{gt} \right) + \sum_{j \in \mathcal{B}} \sum_{ t \in \mathcal{T} } \left( L^{penalty} L_{jt} + O^{penalty} O_{jt} \right)
\end{split}
\\ 
\text{s.t.} \qquad & \parbox{3cm}{\it \small State Equations}  
\begin{dcases}
x_{gt} - x_{gt-1}  = s_{gt} - z_{gt},  & \forall g \in \mathcal{G}, \, t \in \mathcal{T}, \label{con:state_transition}
\end{dcases}  \\ 
%
& \parbox{3cm}{\it \small Minimum Up/Downtime Restrictions}
\begin{dcases}
 \sum_{j=t-UT_g+1}^{t-1} s_{gt} \leq x_{gt}, & \forall g \in \mathcal{G},\, t \in \mathcal{T}, 
%\label{con:min_uptime} 
\\ 
 \sum_{j=t-DT_g}^{t} s_{gt} \leq 1-x_{gt}, & \forall g \in \mathcal{G}, \, t \in \mathcal{T}, 
%\label{con:min_downtime}
\label{con:min_updown}
\end{dcases}
\\
%
& \parbox{3cm}{\it \small Generation Limits}
\begin{dcases}
G_g^{\min} x_{gt} \leq G_{gt} \leq G_g^{\max} x_{gt}, & \forall g \in \mathcal{G}, \, t \in \mathcal{T}, \label{con:gen_limits} 
\end{dcases} \\
% 
& \parbox{3cm}{\it \small Ramping Limits} 
\begin{dcases} 
 \Delta G_g^{\min} \leq G_{gt} - G_{gt-1} \leq \Delta G_g^{\max}, & \forall g \in \mathcal{G}, \, t \in \mathcal{T}, \label{con:ramping} 
\\ 
\end{dcases} \\
%
& \parbox{3cm}{\it \small Flow Limits} 
\begin{dcases} 
F^{\min}_{ij,t} \leq F_{ij,t} \leq F_{ij,t}^{\max}, & \forall (i,j) \in \mathcal{L}, t \in \mathcal{T}, \\ 
\end{dcases} 
\\
& \parbox{3cm}{\it \small Flow Balance} 
\begin{dcases} 
\sum_{i \in \mathcal{B} : (i,j) \in \mathcal{L}} F_{ij,t} - \sum_{i \in \mathcal{B} : (j,i) \in \mathcal{L}} F_{ji,t} + \sum_{g \in \mathcal{G}_j} G_{gt} + L_{jt} - O_{jt} = D_{jt}, & j \in \mathcal{B}, t \in \mathcal{T}, \\ 
\end{dcases} 
\\
& \parbox{3cm}{\it \small Power Flow \newline Approximation} 
\begin{dcases} 
F_{ij,t} = B_{ij} (\theta_{it} - \theta_{jt}), & \forall (i,j)\in\mathcal{L}, t \in \mathcal{T}, 
\end{dcases} 
\\
& \parbox{3cm}{\it \small Integrality Restrictions \& Variable \newline Bounds} 
\begin{dcases} 
x_{gt}, \, s_{gt}, \, z_{gt} \in \{ 0 , 1 \}, & \forall g \in \mathcal{G}, \, t \in \mathcal{T},
\\ 
G_{gt} \geq 0, & \forall g \in \mathcal{G}, \, t \in \mathcal{T},  
\\ 
-\pi \leq \theta_{jt} \leq \pi, ~ L_{jt} \geq 0, \, O_{jt} \geq 0 & \forall j \in \mathcal{B}, \, t \in \mathcal{T}, 
\end{dcases} 
\end{align}
\end{subequations}

In deterministic UC formulations, alternative representations of the above constraints were used to gain computational performance. For brevity, such representations are omitted from this paper. 

\subsection{Economic Dispatch}



\newpage
{\bf Backdrop}: 
\begin{enumerate}
\item {\it ``Scheduling and Pricing for Expected Ramp Capability in Real-Time Power Markets''}: Variability is easier to accommodate than uncertainty. Since the changes are known, it can be scheduled for as long as the flexibility is available and resources are willing to provide it. It is possible that the capabilities of the supply mix are insufficient, and that even if the ramp required is known, the aggregate response speed is not capable of meeting it. When there is sufficient flexibility, the majority of expected variability that occurs on the system can usually be met explicitly through the previously discussed scheduling models. There are a few exceptions. First, it is possible that the variability is occurring at faster resolution than the scheduling model, and not enough information is available to meet this variability explicitly. Second, it is possible that the expected variability occurring in the future can affect the decisions that must be made now, and that information
is beyond the scheduling horizon and not used within the scheduling model. 
\end{enumerate}

\newpage
\begin{thebibliography}{9}

\bibitem{AEO2016}
    \emph{``Annual Energy Outlook''}, 2016, U.S. Energy Information Administration, DOE/EIA-0383.

\bibitem{Gang16a}
	H. Gangammanavar, S. Sen and V. M. Zavala, \emph{``Stochastic Optimization of Sub-Hourly Economic Dispatch With Wind Energy,''} in IEEE Transactions on Power Systems, vol. 31, no. 2, pp. 949-959, March 2016.

\bibitem{Gang16b} H. Gangammanavar and S. Sen, \emph{``Two-scale Stochastic Optimization for Controlling
Distributed Storage Devices,''} accepted for publication in IEEE Transactions on Smart Grid, 2016.

\bibitem{SB350}  California Legislative Information, \emph{``SB-350 Clean Energy and Pollution Reduction Act of 2015''} , Sacramento, CA, October 7, 2015.

\bibitem{Duck16} \emph{``California ISO - Fast Facts''} available at \href{https://www.caiso.com/Documents/FlexibleResourcesHelpRenewables_FastFacts.pdf}{https://www.caiso.com}.

\bibitem{FracMoo16} \emph{``Flexible resource adequacy criteria and must offer obligations''}, available at \href{https://www.caiso.com/informed/Pages/StakeholderProcesses/FlexibleResourceAdequacyCriteria-MustOfferObligations.aspx} {https://www.caiso.com}.
%\bibitem{FracMoo16} \emph{``​Flexible resource adequacy criteria and must offer obligations''}, available at \href{https://www.caiso.com/informed/Pages/StakeholderProcesses/FlexibleResourceAdequacyCriteria-MustOfferObligations.aspx} {https://www.caiso.com}.

\bibitem{Chri99} R. D. Christie, \emph{``Power Systems Test Case Archive''}, available at \href{http://www.ee.washington.edu/research/pstca/} {http://www.ee.washington.edu/research/pstca/}, 1999.

\bibitem{Pott08} C. W. Potter, et al., \emph{``Creating the Dataset for the Western Wind and Solar Integration Study ({U}.{S}.{A}.)''}, Wind Engineering, vol. 32, no. 4, pp. 325-338, 2008.

\bibitem{Higl94} J.L. Higle, and S. Sen, \emph{``Finite master programs in regularized stochastic decomposition''}, Mathematical Programming, vol. 67, pp. 143-168, 1994.

\bibitem{Olso15} A. Olson et al., \emph{``Halfway There: Can California Achieve a 50\% Renewable Grid?},'' in IEEE Power and Energy Magazine, vol. 13, no. 4, pp. 41-52, July-Aug. 2015.

\end{thebibliography}

\newpage

\end{document}
